---
title : "Dynamic Data Analysis"
Author: "Hawnaz O. Najmalddin"
date  : "29.11.2023" 
output: html_notebook
---



# 01) Description: 
this is a dynamic data analysis tool to analyse the iv-ip-ctl-cisplatin H&E tissues. 


`input`
the input folder must contain normalized measurements of different classed from QuPath, named properly, the name of the classes must be written at the end of the file name separated by "_". 

`output`
the output folder will contain rds and csv outputs. 

plot: 
this folder will show a dynamic plot of Average and SEM positive cell proportion in the ROIs in per class, shown ascending. 


```{r LoadLibs, echo=FALSE, message=F, warning=F}
myLibs <- c("dplyr", "data.table")
lapply( myLibs, "require", character.only =T)

# location of raw data: 

dirIn <- file.path("outputs", "NormalizedOCLResults")

```





```{r}
# 2) Load datasets dynamically

# list all files: 
files <- list.files(dirIn, full.names = TRUE)

# function to extract numeric identifier from file name
get_id <- function(file) {
  # extract numeric identifier from the beginning of the file name
  id <- as.numeric(gsub("[^0-9]", "", basename(file)))
  return(id)
}

# iterate over the list of files and load data frames dynamically
dfs <- lapply(files, function(file) {
  # load data frame
  df <- data.table::fread(input = file)
  # get the numeric identifier from the file name
  id <- get_id(file)
  # assign a dynamic name to the data frame based on the identifier and file type
  assign(paste0("df_", id, "_", tools::file_path_sans_ext(basename(file))), df, envir = .GlobalEnv)
})

# clean up unused variables
rm(files)

# Print the names of the loaded data frames
print(ls(pattern = "^df_"))


```



```{r}
gc(full = T)
```

# 02) Apply column remover: 

```{r}
# Load the dplyr package if not already loaded
if (!requireNamespace("dplyr", quietly = TRUE)) {
  install.packages("dplyr")
}
library(dplyr)

# Define the updated colRemover function using dplyr
# Define the updated colRemover function
colRemover <- function(x) {
  x <- x[, -c(1,2,4:8), drop = FALSE]   # this depends on the csv files; must be updated. 
  return(x)
}


gc(full = TRUE)

```

```{r}
rm(dfs); gc(full = T)

```


# 03) Remove unwanted columns by idx.

```{r}
# Loop through all data frames in the global environment
for (df_name in ls(pattern = "^df_")) {
  # Get the data frame using get() function
  df <- get(df_name)
  
  # Remove columns by name using the function
  df <- colRemover(df)
  
  # Assign the modified data frame back to the global environment
  assign(df_name, df, envir = .GlobalEnv)
}

# Print a message indicating completion
cat("Columns removed from all data frames.\n")
```



# 04) label encoding dataframes: 

```{r}
df_list <- ls(pattern = "df_")
# 
# 
# # Use lapply to modify each dataframe in-place
# lapply(df_list, function(df_name) {
#   # Assign the dataframe to the variable
#   df <- get(df_name)
#   
#   # Add a new column 'lbl' filled with NA values
#   df$lbl <- NA
#   
#   # Update the dataframe in the Global environment
#   assign(df_name, df, envir = .GlobalEnv)
# })
# 



lapply(df_list, function(df_name) {
  # Assign the dataframe to the variable
  df <- get(df_name)
  
  # Add a new column 'lbl' filled with the corresponding dataframe name
  df$lbl <- df_name
  
  # Update the dataframe in the Global environment
  assign(df_name, df, envir = .GlobalEnv)
})



```


```{r}
rm(df_list, df_name, df)
```



```{r}
# List all dataframes in the Global Environment
all_dataframes <- ls(pattern = "^df_")

# Create final_A by combining all dataframes
final_A <- do.call(rbind, lapply(all_dataframes, get))

# Remove all other dataframes
rm(list = all_dataframes)
gc(full = T)
```


```{r}
# saveRDS(object = final_A, file = file.path("outputs/labeled_cbindDataSet.rds"))
```




# 05) Read Processed Dataset: 

```{r}
library(dplyr)
df <- readRDS("outputs/labeled_cbindDataSet.rds")
```

```{r}
df$lbl %>% as.factor() %>% levels()
```





```{r}
# Convert "lbl" column to a factor
df$lbl <- as.factor(df$lbl)

# Create a contingency table
contingency_table <- table(df$lbl, df$Name)

# Print the contingency table
print(contingency_table)

```


## read CSV: 

```{r}

df <- data.table::fread("outputs/FinalOutput.csv")
df$Class <- as.factor(df$Class)
df$Count <- as.numeric(df$Count)
df$Rep <- as.factor(df$Rep)

```





```{r}
library(data.table)

# Assuming your data frame is named 'df'
# Convert it to a data.table
dt <- data.table(df)

# Use dcast to reshape the data
result <- dcast(dt, Name ~ Class, value.var = "Count")

# Extract the 'Class' from the 'Name' column
result[, Class := sub('.*_(.*?)$', '\\1', Name)]

# Calculate the Percentage
result[, Percentage := (Positive / (Positive + Negative)) * 100]

# Print the updated result
print(result)



```


```{r}
library(dplyr)

# Assuming 'result' is your current dataframe
Export_df <- result %>%
  group_by(Class) %>%
  summarize(
    Averages = mean(Percentage),
    N = n(),
    STDEV = sd(Percentage),
    SEM = sd(Percentage) / sqrt(N)
  )

# Print the final dataframe
print(Export_df)

```


```{r}
# Load necessary libraries
library(ggplot2)
library(stats)  # Load 'stats' after 'ggplot2' to avoid conflicts
library(plotly)

# Arrange the data by Averages in ascending order
Export_df <- Export_df[order(Export_df$Averages), ]

# Create a column chart with whiskers using ggplot2
plot <- ggplot(Export_df, aes(x = Class, y = Averages)) +
  geom_col() +
  geom_errorbar(aes(ymin = Averages - SEM, ymax = Averages + SEM), width = 0.2) +
  labs(title = "Column Chart with Whiskers",
       x = "Class",
       y = "Averages") +
  theme_minimal()

# Convert ggplot chart to plotly
plotly_chart <- plotly::ggplotly(plot)  # Specify 'plotly::' to avoid conflicts

# Display the interactive plot
plotly_chart

```





